A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Figures

### 2021-08-01 writeUp/supplementary_materials.tex
updated som

\section{Power analysis of MRR studies}
Consider an ideal MRR experiment, where the assumptions of no emigration and harmless marking are fully satisfied. With these data, how accurately can we estimate mosquito lifespan? To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis). Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case, and determined how statistical power to estimate lifespan depends on these two parameters (Fig. \ref{fig:mrr_mcPowerAnalysis}). In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model (eqn. \ref{eq:NB}) where the recapture probability parameters ($\psi$ and $\kappa$) were the averages of those estimated for the actual data.

Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Fig. \ref{fig:mrr_mcPowerAnalysis}A). However once a study length is much longer than the lifespan of the majority of mosquitoes, the predictive power cannot be improved by extending the study duration. The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Fig. \ref{fig:mrr_mcPowerAnalysis}B). For the parameters we use, there are significant gains in accuracy from releasing $1,000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis. For each parameter set we generated 400 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood. The error bars show the standard deviation in the prediction error.}
	\label{fig:mrr_mcPowerAnalysis}
\end{figure}

>>>>

\section{Power analysis of MRR studies}
Consider an ideal MRR experiment, where mosquitoes do not migrate out of the study area and marking is harmless. With these data, how accurately can we estimate mosquito lifespan? To address this question we generated artificial MRR data and attempted to estimate the (known) parameters by maximum likelihood (a `Monte Carlo' analysis). Specifically, we simulated releases of $N$ mosquitoes that are monitored for $m$ days in each case (with collections occurring on each day), and determined how the errors in estimating lifespan depended on these two parameters. In order to focus on estimating lifespan, we assumed the mosquitoes experience constant mortality, and recaptures follow the negative binomial sampling model where the recapture probability parameters ($\psi=0.03$ and $\kappa=3.18$) were the averages of the posterior draws across all species in the species-level model. We assume that all parameters: the true lifespan, $\psi$ and $\kappa$ were unknown and were estimated via maximum likelihood. Here, we investigated two lifespans: 5 days and 10 days indicative of estimates obtained in our analyses.

Unsurprisingly, the error in predicting lifespan declines as the duration of study is increased (Fig. \ref{fig:mrr_mcPowerAnalysis}A). However once a study length is much longer than the lifespan of the majority of mosquitoes, the predictive power cannot be improved by extending the study duration. The effect of increasing release size (while holding study length constant) is similarly a case of diminishing returns (Fig. \ref{fig:mrr_mcPowerAnalysis}B). For the parameters we use, there are significant gains in accuracy from releasing $1000$ rather than $100$ mosquitoes, but modest gains from releasing $10,000$ rather than $1000$ mosquitoes.

Need to include information about power analysis lifespan!

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis.pdf}}
	\caption{\textbf{MRR: Monte Carlo analysis of errors in predicting mean lifespan.} This shows the Root Mean Square Error (RMSE) in predicting mean lifespan as a function of study length (left) and number of released mosquitoes (right) in a Monte Carlo analysis. For each parameter set we generated 200 simulated data series using the negative binomial sampling model with an exponential survival function as described in the text, and estimated the mean lifespan using maximum likelihood. The error bars show the standard deviation in the prediction error.}
	\label{fig:mrr_mcPowerAnalysis}
\end{figure}
